---
title: "Package of the Month - DT"
author: "Christopher Teixeira"
date: 2015-11-10T00:00:00-05:00
categories: ["R"]
tags: ["R", "POTM", "DT", "visualization"]
---

Sometimes you have the need to email someone a dataset for them to analyze. In the past, I always sent an Excel workbook with a bit of formatting. This always takes a bit of time, and never looks as pretty as you want it to. Plus, Excel workbooks just don't work well on cell phones. I've found a nice package that is a wrapper for the DataTables plug-in for jQuery. This package allows me to email out a html file for people to review and analyze as they see fit. Below are just a sample of options I've found handy.

First, let's load in some data and view the default look of the table.

```{r first}
library(Lahman)
library(DT)
data(Master)

datatable(head(Master,n=100))
```

By default, we can search the table and sort individual columns. Let's change the page length to 20 and remove row names.

```{r second}
datatable(head(Master,n=100),
          rownames=F,
          options=list(
            pageLength=20
          ))
```

Sometimes you want to set the default sort order. Here we sort the table by weight descending.

```{r third}
datatable(head(Master,n=100),
          options=list(
            order = list(list(16, 'desc'))
          ))
```

One really cool extension for DataTables allows you to drag and drop columns into a custom order as you explore the data. Combine this with the ability to hide columns really helps you customize a look for the table.

```{r fourth}
dt <- datatable(head(Master,n=100),
          rownames = F,
          extensions = c('ColReorder','Buttons'),
          # extensions = 'Buttons', options = list(dom = 'Bfrtip', )
          options = list(
            pageLength = 20,
            order = list(list(20, 'desc')),
            dom = 'RC<"clear">lfrtip',
            buttons = I('colvis'),
            columnDefs = list(list(visible=F, targets=c(seq(1,14),16,17,seq(21,25))))
          ))

saveWidget(dt,"DT example.html")
```

Note, to get this to look somewhat nice on my blog, I automatically hide several columns.